---
title: "Introduction to the ctrailsgov Package"
author: Taylor Arnold and Michael J Kane
output:
  rmarkdown::html_vignette:
    css: "note-style.css"
  vignette: >
    %\VignetteIndexEntry{Introduction to the ctrailsgov Package}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

This vignette gives a very brief overview of the current package. To start,
we load the package into R.

```{r, message=FALSE}
library(ctrialsgov)
```

In the next few sections, we see how to setup the data set, query it, and then
visualize the output.

## Create the Data

Before querying the ClinicalTrials.gov data, we need to load a pre-processed
version of the data into R. There are three ways to do this. If you have
installed a copy of the data set locally into PostGRES, the data can be
created from scratch with the following block of code (it will take a couple
of minutes to finish):

```{r, eval=FALSE}
library(DBI)
library(RPostgreSQL)

drv <- dbDriver('PostgreSQL')
con <- DBI::dbConnect(drv, dbname="aact")
ctgov_create_data(con)
```

Alternatively, we can download a static version of the data from GitHub and
load this into R without needing the setup a local version of the database.
This will be cached locally so that it can be re-loaded without downloading
each time. To download and load this data, use the following:

```{r, eval=FALSE}
ctgov_load_cache()
```

Finally, we can load a small sample dataset (2% of the total) that is included
with the package itself using the following:

```{r}
ctgov_load_sample()
```

This is the version of the data that is used in most of the tests, examples,
and in this vignette.

## Querying the Data

The primary function for querying the dataset is called `ctgov_query`. It can
be called after using any of the functions in the previous section. Here
are a few examples of how the function works. We will see a few examples here;
see the help pages for a complete list of options.

There are a number of fields in the data that use exact matches of categories.
Here, for example, we find the interventional studies:

```{r}
ctgov_query(study_type = "Interventional")
```

Or, all of the interventional studies that have a primary industry sponsor:

```{r}
ctgov_query(study_type = "Interventional", sponsor_type = "INDUSTRY")
```

A few fields have continuous values that can be searched by giving a vector
with two values. The results return any values that fall between the lower
bound (first value) and the upper bound (second value). Here, we find the
studies that have between 40 and 42 patients enrolled in them:

```{r}
ctgov_query(enrollment_range = c(40, 42))
```

Setting one end of the range to missing avoids searching for that end of the
range. For example, the following finds any studies with 1000 or more patients.

```{r}
ctgov_query(enrollment_range = c(1000, NA))
```

Similarly, we can give a range of dates. These are given in the form of strings
as "YYYY-MM-DD":

```{r}
ctgov_query(date_range = c("2020-01-01", "2020-02-01"))
```

Finally, we can also search free text fields using keywords. The following for
example finds and study that includes the phrase "lung cancer" (ignoring
case) in the description field:

```{r}
ctgov_query(description_kw = "lung cancer")
```

We can search two terms at once as well, by default it finds things that match
at least one of the terms:

```{r}
ctgov_query(description_kw = c("lung cancer", "colon cancer"))
```

But the `match_all` flag can be set to search for both terms at the same time
(here, that returns no matches):

```{r}
ctgov_query(description_kw = c("lung cancer", "colon cancer"), match_all = TRUE)
```

Other keyword fields include `official_title_kw`, `source_kw` and `criteria_kw`.

Any of the options can be combined as needed.

```{r}
ctgov_query(
  description_kw = "cancer",
  enrollment_range = c(100, 200),
  date_range = c("2019-01-01", "2020-02-01")
)
```

Finally, we can also pass a current version of the data set to the query
function, rather than starting with the full data set. This is useful when
you want to combine queries in a more complex way. For example, this is
equivalent to the above:

```{r, message=FALSE}
library(dplyr)

ctgov_query() %>%
  ctgov_query(description_kw = "cancer") %>%
  ctgov_query(enrollment_range = c(100, 200)) %>%
  ctgov_query(date_range = c("2019-01-01", "2020-02-01"))
```

## Visualizing the output

The package also contains a number of tools for visualizing the output.
Here is one example:

```{r}
ctgov_query(
  description_kw = "cancer",
  enrollment_range = c(100, 200),
  date_range = c("2019-01-01", "2020-02-01")
) %>%
  ctgov_plot_timeline() +
    ggplot2::theme_minimal()
```
